### 2019-12-16 report/02-methods.Rmd
Group PCA by relevant part + add temp & DBO to SEM

@@

bluefunc <- colorRampPalette(c("lightyellow", "darkred"))
ordiplot(nmds,type="n")
orditorp(nmds,display="species",col="gray",air=0.01)
orditorp(nmds, display="sites",
  col= bluefunc(5)[findInterval(tps_com_met$richness_avg, seq(2,6))],
  # cex = scale(tps_com_met$biomass_med, center = FALSE),
  air=0.01, pch = "+")
```

>>>>

#### Separate Med and CV pressure

```{r pca-med-cv, fig.cap="PCA with Avg or CV variables"}
hab_med <- habitat_pressure %>%
  select_at(vars(matches("med|mean")))
hab_cv <- habitat_pressure %>%
  select_at(vars(matches("_cv"))) %>%
  select(-width_river_cv, -avg_depth_station_cv, -nitrates_cv, -phosphore_cv)

plot_pca <- function (.data, title = NULL) {

  if (is.null(title)){
  title <- paste0("PCA - ",
      ifelse(any(!is.na(str_match(names(.data), "_cv"))), "CV",
      "Median"))
  }

  pca_hab <- dudi.pca(as.data.frame(na.omit(.data)), scannf = FALSE, nf = 3, center = TRUE, scale = TRUE) 
  eig_plot <- fviz_eig(pca_hab)
  pca_plot <- fviz_pca_var(pca_hab,
    axes = c(1, 2),
    col.var = "contrib", # Color by contributions to the PC
    gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
    title = title,
    repel = TRUE # Avoid text overlapping
  )
  pca_plot12 <- ggdraw(pca_plot) +
    draw_plot(eig_plot, x = .7, y = .95, hjust = 1, vjust = 1, width = 0.17, height = 0.25)
  pca_plot12
  }
plot_pca_med_cv <- map(list(hab_med, hab_cv), plot_pca)
plot_grid(plotlist = plot_pca_med_cv, labels = "AUTO")
```

#### Separate according to physical habitat, climate, enrichment

```{r pca-evt}
phys_var <- c( "width_river_mean", "avg_depth_station_mean",
   "slope", "alt", "d_source", "strahler")
#"width_river_cv", "avg_depth_station_cv", "under_bank_med", "hole_pit_med",
#"long", "lat","logjam_stumps_med", "edge_vg_med",
#"rock_shelter_med", "aqua_vg_shelter_med", "shade_med", "sinuosite_med",
climate_var <- c("flow_med", "temperature_med", "flow_cv", "temperature_cv") 
pressure_var <- c( "fungicides_cv", "herbicides_cv",
  "insecticides_cv",  "fungicides_med",
  "herbicides_med", "insecticides_med")
#, "nitrates_med", "phosphore_med","nitrates_cv", "phosphore_cv","DBO_med", "DBO_cv"
hab_physical <- habitat_pressure[, names(habitat_pressure) %in% phys_var]
hab_climate <- habitat_pressure[, names(habitat_pressure) %in% climate_var]
hab_pressure <- habitat_pressure[, names(habitat_pressure) %in% pressure_var]

p_pca_clim_phy_press <- map2(
  list(hab_physical, hab_climate, hab_pressure),
  c("Physical", "Climate", "Pressure"),
  plot_pca)
plot_grid(plotlist = p_pca_clim_phy_press, labels = "AUTO")
```

@@

bluefunc <- colorRampPalette(c("lightyellow", "darkred"))
ordiplot(nmds,type="n")
orditorp(nmds,display="species",col="gray",air=0.01)
orditorp(nmds, display="sites",
  col= bluefunc(15)[findInterval(tps_com_met$richness_avg, seq(1,15))],
   cex = findInterval(log10(tps_com_met$biomass_med), seq(1, 4, 1))/2,
  air=0.01, pch = "+", labels = "+")

### 2019-12-16 report/05-stability.Rmd
Group PCA by relevant part + add temp & DBO to SEM

@@

```{r}
biomass_sem <- temporal_community_metrics %>%
  select(station, biomass_stab, biomass_med, richness_med, pielou_med,
  beta_bin_c) %>%
  left_join(select(synchrony, station, synchrony, cv_sp), by = "station") %>%
  rename(sync = synchrony,
    piel = pielou_med,
    prod = biomass_med 
  )
habitat_sem <- habitat_pressure %>% 
  select(station, alt, width_river_mean) %>%
  rename(width = width_river_mean)
net_sem <- temporal_network_metrics %>%
  select(station, connectance_corrected_med, w_trph_lvl_avg_med) %>%
  rename(c_c = connectance_corrected_med,
    t_lvl = w_trph_lvl_avg_med)
composition_axis <- nmds$points %>%
  as_tibble %>%
  mutate(station = as.integer(dimnames(nmds$points)[[1]]))

dsem <- biomass_sem  %>%
  mutate(log_rich = log10(richness_med)) %>%
  left_join(net_sem, by = "station") %>%
  left_join(habitat_sem, by = "station") %>%
  left_join(st_basin, by = "station") %>%
  left_join(composition_axis, by = "station") %>%
  na.omit()

@@

#log_stab ~ log_sync + log_cv_sp
lav_formula <- "
  log_sync ~ log_rich + piel + alt + c_c + t_lvl + beta_bin_c
  log_cv_sp ~ log_rich + piel + c_c + t_lvl + beta_bin_c
  log_rich ~ alt + width 
  piel ~ alt + width
  c_c ~ alt + width
  t_lvl ~ alt + width
  beta_bin_c ~ alt + width
  log_cv_sp ~~ log_sync
  "
lav_mod <- sem(lav_formula, lav_data)
standardizedsolution(lav_mod) %>%
  kable() %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "400px")
```

@@

```{r, fig.cap = "The basic SEM"}
  corsem <- psem(
    lme(log_sync ~ log_rich + piel + alt + c_c + t_lvl + beta_bin_c, random = ~ 1 | basin, data = lav_data),
    lme(log_cv_sp ~ log_rich + piel + c_c + t_lvl + beta_bin_c, random = ~ 1 | basin, data = lav_data),
    lme(log_rich ~ alt + width, random = ~ 1 | basin, data = lav_data),
    lme(piel ~ alt + width, random = ~ 1 | basin, data = lav_data),
    lme(c_c ~ alt + width, random = ~ 1 | basin, data = lav_data),
    lme(t_lvl ~ alt + width, random = ~ 1 | basin, data = lav_data),
    lme(beta_bin_c ~ alt + width, random = ~ 1 | basin, data = lav_data),
    lm(log_stab ~ log_cv_sp + log_sync, data = lav_data)
  )
output <- summary(corsem, .progressBar = F)
output$coefficients %>%
  kable() %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "400px")

source(mypath("R", "plot_methods.R"))
ggpiecewisesem(output)
```

```{r}
compute_psem <- function(.data) {
  corsem <- psem(
    lme(log_sync ~ log_rich + piel + alt + c_c + t_lvl + beta_bin_c + MDS2, random = ~ 1 | basin, data = .data),
    lme(log_cv_sp ~ log_rich + piel + c_c + t_lvl + beta_bin_c + MDS2, random = ~ 1 | basin, data = .data),
    lme(log_rich ~ alt + width, random = ~ 1 | basin, data = .data),
    lme(MDS2 ~ alt + width, random = ~ 1 | basin, data = .data),
    lme(piel ~ alt + width, random = ~ 1 | basin, data = .data),
    lme(c_c ~ alt + width, random = ~ 1 | basin, data = .data),
    lme(t_lvl ~ alt + width, random = ~ 1 | basin, data = .data),
    lme(beta_bin_c ~ alt + width, random = ~ 1 | basin, data = .data),
    lm(log_stab ~ log_cv_sp + log_sync, data = .data)
  )
  output <- summary(corsem, .progressBar = F)

@@

>>>>

```{r}
myload(nmds, dir = mypath("report"))

@@

```{r}
biomass_sem <- temporal_community_metrics %>%
  select(station, biomass_stab, biomass_med, richness_med, pielou_med,
  beta_bin_c) %>%
  left_join(select(synchrony, station, synchrony, cv_sp), by = "station") %>%
  rename(sync = synchrony,
    piel = pielou_med,
    prod = biomass_med 
  )
habitat_sem <- habitat_pressure %>% 
  select(station, lat, alt, width_river_mean, DBO_med, DBO_cv, flow_med, flow_cv,
    temperature_med, temperature_cv) %>%
  rename(width = width_river_mean)
net_sem <- temporal_network_metrics %>%
  select(station, connectance_corrected_med, w_trph_lvl_avg_med) %>%
  rename(c_c = connectance_corrected_med,
    t_lvl = w_trph_lvl_avg_med)
composition_axis <- nmds$points %>%
  as_tibble %>%
  mutate(station = as.integer(dimnames(nmds$points)[[1]]))

dsem <- biomass_sem  %>%
  mutate(log_rich = log10(richness_med)) %>%
  left_join(net_sem, by = "station") %>%
  left_join(habitat_sem, by = "station") %>%
  left_join(st_basin, by = "station") %>%
  #left_join(composition_axis, by = "station") %>%
  na.omit()

@@

#log_stab ~ log_sync + log_cv_sp
lav_formula <- "
  log_sync ~ log_rich + piel + alt + c_c + t_lvl + beta_bin_c
  log_cv_sp ~ log_rich + piel + c_c + t_lvl + beta_bin_c
  log_rich ~ alt + width 
  piel ~ alt + width
  c_c ~ alt + width
  t_lvl ~ alt + width
  beta_bin_c ~ alt + width
  log_cv_sp ~~ log_sync
  "
lav_mod <- sem(lav_formula, lav_data)

standardizedsolution(lav_mod) %>%
  kable() %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "400px")
```

@@

```{r, fig.cap = "The basic SEM"}
build_stab_sem <- function (evt = NULL, com = NULL, stab_comp = NULL, stab = NULL,  .data = NULL) {

  rs <- map(list(evt = evt, com = com, stab_comp = stab_comp), ~paste(.x, collapse = " + "))
  stab_eq <- paste(stab, rs[["stab_comp"]], sep = " ~ ")
  stab_comp_eq <- map_chr(stab_comp, ~ paste(.x, rs[["com"]], sep = " ~ "))
  com_eq <- map_chr(com, ~paste(.x, rs[["evt"]], sep = " ~ "))

  formulas <- map(c(stab_eq, stab_comp_eq, com_eq), as.formula)

  output <- vector("list", length = length(formulas))

  output[[1]] <- lm(as.formula(stab_eq), data = .data)

  lmer_models <- map(formulas[2:length(formulas)],~lme(fixed = .x, random = ~ 1 | basin, data = .data))
  lmer_models_c <- map2(lmer_models, formulas[2:length(formulas)],
    function (.model, .formula){
      .model$call$fixed <- as.call((.formula))
      return(.model)
  })

  output[2:length(output)] <- lmer_models_c
  return(output)

}

#undebug(build_stab_sem)
corsem <- build_stab_sem(
  evt = c("alt", "width", "DBO_med", "temperature_med"),
  com = c("log_rich", "piel", "c_c", "t_lvl", "beta_bin_c"),
  stab_comp = c("log_sync", "log_cv_sp"),
  stab = c("log_stab"),
  .data = lav_data)

output <- summary(as.psem(corsem), .progressBar = F)
output$coefficients %>%
  kable() %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "400px")

source(mypath("R", "plot_methods.R"))
ggpiecewisesem(output)
```
```{r}

car::vif(lme(log_sync ~ log_rich + piel + alt + c_c + t_lvl + beta_bin_c, random = ~ 1 | basin, data = lav_data))
car::vif(lme(log_sync ~ alt + width + temperature_med + DBO_med + flow_med, random = ~ 1 | basin, data = lav_data))
```

```{r}
compute_psem <- function(.data) {
  corsem <- psem(
    lme(temperature_med ~ alt, random = ~ 1 | basin, data = .data),
    lme(log_rich ~ alt + width + temperature_med + DBO_med, random = ~ 1 | basin, data = .data),
    lme(piel ~ alt + width + temperature_med + DBO_med, random = ~ 1 | basin, data = .data),
    lme(c_c ~ alt + width + temperature_med + DBO_med, random = ~ 1 | basin, data = .data),
    lme(t_lvl ~ alt + width + temperature_med + DBO_med, random = ~ 1 | basin, data = .data),
    lme(beta_bin_c ~ alt + width + temperature_med + DBO_med, random = ~ 1 | basin, data = .data),
    lme(log_sync ~ log_rich + piel + alt + c_c + t_lvl + beta_bin_c,
      random = ~ 1 | basin, data = .data),
    lme(log_cv_sp ~ log_rich + piel + c_c + t_lvl +
    beta_bin_c, random = ~ 1 | basin, data = .data),
    lm(log_stab ~ log_cv_sp + log_sync, data = .data)
  )
  output <- summary(corsem, .progressBar = F)

@@

- Main results:
  - Richness destabilizes communities 
  - Turnover destabilizes communities, latter increased by enrichment
  - Trophic level decreased CV but increase species synchrony!:
  - Connectance!
